### 2023-08-03 manuscript/supplementary_methods.Rmd

---
output: pdf_document
bibliography: references.bib
---

```{r setup,include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE)
```

# Supplementary materials and methods

## Sample collection and extraction protocols

### *Brachypodium distachyon* tissue collection and extraction

Plants of *Brachypodium distachyon* ecotypes ABR1, ABR5, ABR6 and Bd21 were grown for 21 days in environmentally controlled growth rooms (Polysec, R. J. Hicks Refrigeration, Aberystwyth, UK) at 23 ^$\circ$^C under a 16 hour light period. 
A total of 15 plants of each ecotype were harvested by detaching the 2nd fully developed leaf from the base of the plant and dissecting the middle 4 cm segment. 
Each sample was then placed in a 2 mL Eppendorf tube containing a 4mm steel ball bearing and snap frozen in liquid nitrogen.
Samples were stored at -80 ^$\circ$^C until extraction. 
All harvesting was conducted in the growth room to avoid environmental fluctuations.

The frozen samples were milled using a Retsch 301 Mixer Mill at 30 Hz for 30 seconds and then placed on crushed ice. 
700 $\mu$L of pre-chilled extraction solvent (CHCl~3~:MeOH:H~2~O; 1:2.5:1; v:v:v) was immediately added and suspended by vortexing (Scientific Industries Vortex Genie-2). 
Samples were placed on an orbital shaker (FATSN002, Favorgen Biotech Corp) for 20 minutes at 1,400 rpm and a temperature of 4 ^$\circ$^C. 
After shaking, the samples were centrifuged (EBA 12R, Hettich) at 13,000 rpm for 6 minutes at 0 ^$\circ$^C. 
The supernatant of each sample was then transferred to a new 2 mL Eppendorf tube and the pellet was discarded. 
The samples were stored at -80 ^$\circ$^C until MS analysis.

### Human urine extraction and addition of chemical standards

Mixed gender, unfiltered pooled human urine (BioIVT, Catalog# HUMANURINEPNN, Lot# 0393918) from healthy donors was extracted  by adding 500 $\mu$L of pre-chilled MeOH (Extraction Grade, Fisher Scientific) and vortexing. 
Samples were placed on an orbital shaker (FATSM002, Favorgen Biotech Corp) for 20 minutes at 1,400 rpm and 4 ^$\circ$^C in the dark. 
The samples were centrifuged (EBA 12 R, Hettich) at 13,000 rpm for 5 minutes at 4 ^$\circ$^C and 20 $\mu$L of the supernatant was transferred to a clean glass HPLC vial, containing a 200 $\mu$L flat bottom micro insert (Chromacol) and diluted with 80 $\mu$L of H~2~O:MeOH (3:7) directly in the vial.

The `r tar_read(spiked_urine_compound_info) %>% nrow()` chemical standards were dissolved in a solvent mixture of methanol, water and formic acid (70%:30%:0.1%), each 4.5 $\mu$g/ml concentration.
700 $\mu$L of urine and 300 $\mu$L of the standards mixture were combined to produce the spiked urine sample.

## MS analysis

FIE-HRMS mass spectra were acquired on an Exactive Orbitrap (ThermoFinnigan, San Jose CA) mass spectrometer coupled to an Accela (ThermoFinnigan, San Jose CA) ultra-performance liquid chromatography (LC) system for the *B. distachyon* samples and the Human urine on a Q-Exactive Orbitrap (ThermoFinnigan, San Jose CA) coupled to a Dionex (ThermoFinnigan, San Jose CA) ultra-performance LC system. 
A sample volume of 20 $\mu$L was injected and delivered to the electrospray ionisation (ESI) source via a 'plug' flow solvent of pre-mixed HPLC grade MeOH (Fisher Scientific) and ultra-pure H~2~O (18.2 $\Omega$) at a ratio of 7:3. 
The initial flow rate was 60 $\mu$L min^-1^ for 0.4 minutes and increased to 200 $\mu$L min^-1^ over 0.8 minutes. 
The flow rate was maintained at 200 $\mu$L min^-1^ for 0.3 minutes then increased to 600 $\mu$L min^-1^ over 1.5 minutes. 
Then the flow rate was then returned to 60 $\mu$L min^-1^ for 0.5 minutes. 
The total gradient time was 3.5 minutes. The capillary temperature and voltage were set at 270 ^$\circ$^C and 37.50 kV respectively.

For the *B. distachyon* samples, mass spectra were acquired using a single scan filter for each of the positive and negative ionisation modes. 
Positive ions were acquired between 55.000 and 1000.000 *m/z*, and  63.000 and 1000.000 *m/z* for negative ions.
Mass spectra for the spiked human urine were acquired with two scan events between 55-280 *m/z* and 270-1200 *m/z* for each of the positive and negative acquisition modes.
For all scan events, the scan rate was 1 Hz with a mass resolution of 100,000. 
The automatic gain control (AGC) target was 5 x10^5^ and the total ion injection time 250 ms.

Each biological matrix was analysed in separate analytical runs.
For the spiked urine, a total of 50 replicate injections were performed.
Three blank injections were preformed prior to the analytical replicate injections, using the solvent mixture used in the sample preparation of the particular matrix. 
For the *B. distachyon* samples, the samples were organised into randomised blocks, with each ecotype class represented by two samples in each block.
Similarly to the spiked urine samples, three blank injections were performed first, followed by three quality control samples (QC) injections.
The QC samples were made up of equal parts of each of the 60 samples in the run.
A QC sample injection then followed each of the random blocks.

C18-UHPLC-HRMS analyses were performed on an Orbitrap Fusion Tribrid mass spectrometer (Thermo Scientific) that was coupled to a Ultimate 3000 liquid chromatography tower (Dionex, Thermo Scientific). 
Mass spectra were acquired using a single scan filter for each of the positive and negative ionisation modes between a range of 100-2000 m/z. For all scan events, the scan rate was 1 Hz with a mass resolution of 100,000. The automatic gain control (AGC) target was 5 x10^5.

A Hypersil Gold reverse phase C18 column (2.1 mm x 150 mm; particle size 1.9 $\mu$m) was used for chromatography, which was maintained at a temperature of 60 $^\circ$C. 
The mobile phases consisted of ultra-pure water - formic acid (100:0.1) (A), LC-MS grade methanol -formic acid (100:0.1) (B) and HPLC grade isopropyl alcohol - LCMS grade methanol (1:1) (C). 
The initial condition was A:B (99:1.0) for 0.5 minutes and the percentage of B increased linearly over 6 minutes, to 60.0%. 
The percentage of B was increased further for another 4 minutes to 100% and held for 2 minutes. 
C was then introduced over 0.1 minutes to 100% and held for 1.4 minutes.
Prior to re-equilibration, 100% C was switched to 100% B in 0.1 minutes and then returned to the starting conditions of A:B (99:1).
This was held for 3.4 minutes giving a total method time was 17 minutes. 
The flow rate was kept at 400 $\mu$L min^-1^ except for between 12.1 and 13.6 minutes where it was reduced to 300 $\mu$L min^-1^.

## Spectral processing

After the MS data were acquired, the raw mass spectra data files (.RAW, ThermoFinnigan) were converted to the universal mass spectrometry open file format, mzML.
The conversion and vendor centroiding were performed using *msconvert* from the Proteowizard tool pack.

The FIE-HRMS fingerprinting data were spectrally processed by performing spectral binning to 0.01 amu using the R package *binneR* v`r packageVersion('binneR')`.
The infusion scans were detected using a total ion count threshold of above 50% of the maximum total ion count across the infusion profile.

For C18-UHPLC-HRMS, spectral processing was performed using the R package *xcms* v`r packageVersion('xcms')`.
Peak detection used the *centWave* algorithm, with the tolerated deviation between consecutive scans set at 8.2 parts per million (ppm), a minimum and maximum peak width of 16 and 43 seconds respectively and a signal to noise threshold of 7. 
The *Obiwarp* algorithm was used for retention time correction with a bin size of 0.7, a gap opening penalty of 2.9 and a gap enlargement penalty of 2.7. 
Peak grouping was performed using peak density, with a bin size of 0.00775 amu, a chromatographic bandwidth of 5 seconds and a minimum group fraction of two thirds. 
The *fillPeaks* method was used to integrate the areas of missing peaks using the default parameters.

## Data pre-treatment

The pre-treatment routines were performed separately for positive and negative ionisation mode features, with the features subsequently combined into a single intensity matrix.

For FIE-HRMS of the spiked urine, *m/z* were first filtered based on an overall occupancy threshold of two thirds.
All missing values were then imputed using random forest imputation. Only *m/z* features with a relative standard deviation (RSD) value below 50% were retained and finally a total ion count (TIC) normalisation was applied to each sample.

For the FIE-HRMS of the *B. distachyon* samples, a median centring batch correction was first performed where the median of each feature for each randomised block of samples were adjusted to the overall median.
The QC samples were used to calculate the RSD of each feature across the run batch.
These samples were first prepared by retaining only features with an occupancy above two thirds, and imputing missing values using random forest imputation.
Features with an RSD below 50% were retained and the QC samples were subsequently removed.
A maximum occupancy feature filter was then applied where features were retained if at least a single ecotype class had an occupancy of above two thirds of the samples.
Class-wise imputation using random forest was performed on features with a class occupancy greater than two thirds. 
Finally, a TIC normalisation was applied to each of the samples.

For the C18-UHPLC-HRMS of the *B. distachyon* samples, these data were pre-treatment as for FIE-HRMS fingerprinting data with the exception that a RSD filtering threshold of 20% was instead applied.

## Performance benchmarking of the *assignments* R package

```{r benchmarking-info}
targets::tar_load(cpu_information)
targets::tar_load(memory_information)
n_reps <- xfun::numbers_to_words(n_reps)
n_correlations <-  n_correlations %>% 
  prettyNum(big.mark=",",
            scientific=FALSE) %>% 
  str_as_written_list()
n_cores <- n_cores %>% 
  str_as_written_list()
```

Processing time was benchmarked for all the combinations of maximum number of correlations (`r n_correlations`) and CPU workers (`r n_cores`).
For each test, the molecular formula assignment routine was performed on the pre-treated spiked urine data matrix with all other parameters as the defaults using the R package *assignments* v`r packageVersion('assignments')`.
Tests were conducted on a Dell high performance computing blade with an `r cpu_information$Value[cpu_information$Type == 'Model name']` model processor with `r cpu_information$Value[cpu_information$Type == 'CPU(s)']` CPUs, `r memory_information$Size[memory_information$Type == 'MemTotal'] %>% as.character()`B of RAM and `r osVersion` operating system.

The R package *rbenchmark* v`r packageVersion('rbenchmark')` was used to measure processing time with `r n_reps` replications of each correlation and CPU workers combination.