### 2023-11-08 manuscript/supplementary_tables.Rmd
removed the matched chemical standards row from table S6 as this is now provided in table 2

---
output: pdf_document
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{xcolor}
  - \newcommand{\beginsupplement}{\setcounter{table}{1}  \renewcommand{\thetable}{S\arabic{table}}}
---

```{r setup,include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE)
```

\beginsupplement

```{r table-s2}
tar_load(brachy_FIE_HRMS_parameters_molecular_formula_assignment)

adducts <-  brachy_FIE_HRMS_parameters_molecular_formula_assignment %>% 
  assignments::adducts() %>% 
  purrr::flatten_chr()

mzAnnotation::adduct_rules() %>% 
  dplyr::filter(Name %in% adducts) %>% 
  dplyr::select(
    Adduct = Name,
    Charge,
    `No. molecules` = xM,
    `Mass difference` = Add
  ) %>% 
  mutate(`Mass difference` = round(`Mass difference`,5)) %>% 
  knitr::kable(caption = 'Information for positive and negative mode adducts used for calculating ionisation products.',
               format = 'latex',
               booktabs = TRUE)
```

```{r table-s3}
mzAnnotation::isotope_rules() %>% 
  dplyr::filter(Isotope %in% c('13C','13C2','18O','34S')) %>% 
  dplyr::select(Isotope,
                `Mass difference` = `Mass Difference`) %>%
  dplyr::mutate(Isotope = c('\\textsuperscript{13}C',
                            '\\textsuperscript{13}C2',
                            '\\textsuperscript{18}O',
                            '\\textsuperscript{34}S')) %>% 
  knitr::kable(caption = 'Isotopic mass difference information.',
               format = 'latex',
               escape = FALSE,
               booktabs = TRUE)
```

```{r table-s4}
transformations <- brachy_FIE_HRMS_parameters_molecular_formula_assignment %>% 
  assignments::transformations()

mzAnnotation::transformation_rules() %>% 
  dplyr::filter(`MF Change` %in% transformations) %>% 
  dplyr::select(
    Transformation = Name,
    `MF change` = `MF Change`,
    `Mass difference` = `Difference`
  ) %>% 
  knitr::kable(caption = 'Transformation mass difference information.',
               format = 'latex',
               booktabs = TRUE)
```

```{r table-s5}
tar_read(processing_time_results) %>% 
  dplyr::select(workers,
                `Number of correlations` = `# correlations`,`time (seconds)`) %>% 
  dplyr::mutate(`time (minutes)` = `time (seconds)` / 60,
                dplyr::across(tidyselect::where(is.numeric),~signif(.x,digits = 3)),
                `Number of correlations` = scales::comma(
                  `Number of correlations`,
                )) %>% 
  dplyr::select(-`time (seconds)`) %>% 
  tidyr::spread(workers,`time (minutes)`) %>% 
  knitr::kable(
    format = 'latex',
    booktabs = TRUE,
    caption = 'Processing time (minutes) results for the R package \\textit{assignments}. Feature \\textit{m/z} were used from spiked urine matrix and the numbers of correlations limited by using the correlations with the highest absolute coefficients up to the designated number. Molecular formula assignment was then performed across the range of numbers of central processing unit (CPU) workers.',
    align = rep('r',4)
  ) %>% 
  kableExtra::kable_styling() %>% 
  add_header_above(c(" ", "CPU workers" = 3))
```

```{r table-s6}
tar_read(assignment_results) %>%
  dplyr::full_join(
    tar_read(assignment_matches) %>% 
      dplyr::mutate(
        matrix = matrix %>% 
          replace(
            matrix == 'standards',
            'standards-mix'
          )
      ), 
    by = c("matrix", "technique")) %>% 
  mutate(
    matrix = paste(matrix,technique)
  ) %>%
  mutate(
    n_assignments = glue::glue(
      '{n_assignments} ({signif(n_assignments/n_features * 100,digits = 3)}\\%)'),
    across(where(is.numeric),~prettyNum(.x,big.mark = ','))) %>%
  select(-technique) %>% 
  dplyr::rename(
    `\\textit{m/z} features` = n_features,
    correlations = n_correlations,
    relationships = n_relationships,
    `adduct and isotopic step assignments` = n_ai_assignments,
    `transformation step assignments` = n_t_assignments,
    `assigned \\textit{m/z}` = n_assignments,
    `molecular formulas (MF)` = n_MFs,
    `KEGG compound matches` = KEGG_compound_matches,
    `KEGG MF matches` = KEGG_MF_matches,
    `matched chemical standards` = matched_standards) %>% 
  tidyr::gather(Variable,Value,-matrix) %>% 
  mutate(
    Value = replace(Value,
                    Value == 'NA',
                    'N/A')) %>% 
  tidyr::spread(matrix,Value) %>% 
  slice(c(1,4,9,2,10,3,8,6,5)) %>%
  set_colnames(
    stringr::str_split_fixed(colnames(.),
                             ' ',2)[,2]) %>% 
  knitr::kable(
    format = 'latex',
    booktabs = TRUE,
    linesep = '',
    escape = FALSE,
    align = c(
      rep('l',1),
      rep('r',3)
    ),
    caption = 'A summary of the molecular formula assignment results for the example sample matrices. The number of matched chemical standards is given as the number of unique molecular formulas.') %>%
  kable_styling() %>%
  add_header_above(c(
    " ",
    "\\\\textit{B. distachyon} leaf" = 2,
    "Human urine" = 1,
    "Standards mix" = 1),
    escape = FALSE)
```