### 2023-11-08 manuscript/additional_files.Rmd
re-number additional files 4-6 inline with the restructuring of section 3.2

# Additional files

Additional file 1.csv - Table S1 -
Chemical identifiers of the compounds spiked into the urine extract.

Additional file 2.pdf - Supplementary materials and methods -
Further details outlining the materials and methods used in this article.

Additional file 3.pdf - Supplementary tables -
Supplementary tables S2, S3, S4, S5, S6 showing adduct, isotopic, transformation, R package benchmarking information and a summary of the molecular formula assignment results for the example sample matrices respectively.

Additional file 4.csv Table S7 -
The *m/z* features directly matched to theoretical ionisation products of the chemical standards in the standards mixed and spiked urine.

Additional file 5.csv Table S8 -
The *m/z* features assigned to chemical standards in the standards mix and spiked urine.

Additional file 6.csv - Table S9 -
The molecular formula assignments for *m/z* across all experimental matrices along with any match KEGG compound IDs.